### 2014-11-10 CVPR/Source/abstract.tex
Reorganized paper tex files

\begin{abstract}
  TODO: Write abstract
  \vspace{4cm}
\end{abstract}