### 2014-11-10 CVPR/Source/related.tex

\section{Related Work}
\label{sec:related}

TODO:
\vfill
\pagebreak
Related work continued
\vspace{15cm}