### 2015-04-13 CVPR/Source/abstract.tex
Updated abstract.

\begin{abstract}
  %Problem:
  3D localization in road scenes from monocular video is an important problem
  for applications in autonomous driving. We propose a probabilistic graphical
  model based framework to estimate 3D localization of traffic participants in
  road scenes.  We use object detections, point tracks, egomotion, estimated
  ground truth, GPS and map information as an input to our system.
  Given this input we compute 6DOF 3D localization of traffic participants
  along with their dimensions (height, width and length).
  We also introduce a novel perspective to the soft occlusion models where a
  region in space is viewed in terms of reflection and transmission
  probability.
  We test and train our model on KITTI dataset and show that our occlusion
  model works better than the baseline method of bounding boxes. We also show
  that our 3D localization results with monocular video input are comparable to
  (Geiger 2014) which uses stereo input.

  % Why monocular
  % Since laser scanners are expensive, we focus on solving the problem through
  % monocular video, GPS and maps as our input. Given the video we extract the 
  % Methodology
  % We formulate the problem of 3D localization in a probabilistic graphical
  % model specifically a factor graph model. We use additional heuristic
  % constraints like collision constraint,

\end{abstract}