### 2015-05-03 presentation/intro.tex

% \begin{frame}{Problem}
%   % Notes: Spend around 30 sec on this slide
%   \centering
%   \begin{figure}
%     \input{Source/problemdef.tex}
%   \end{figure}
% \end{frame}
% This slide doesn't make it clear what problem we are trying to solve?  May be
% the slide that Manmohan made with an example image with cars and detections
% is better

% Should we have introduction slide that highlights our contributions?
% but we should have contributions to highlight them. Contributions are perhaps
% better with positive results but still. I don't think we should have negative
% results as a surprise. (It won't as they would have read the report.)

\begin{frame}{Problem}
  % Need a flowchart that describes the input and desired output of the system
 \tikzset{/tikz/x=0.8cm,/tikz/y=0.8cm}
 \newlength{\units}
 \setlength{\units}{0.8cm}
  \scriptsize{
    \input{Source/problemflowchart.tex}
  }
\end{frame}

% \begin{frame}{Problem}
%   \visible<-1>{
%     \begin{tikzpicture}
%       \node[anchor=south west,inner sep=0] (image) at (0,0)
%       {\includegraphics[width=\textwidth]{graphics/0000000061_bevdown.png}};
%       \fill[color=white] (image.south west) rectangle (image.east);
%     \end{tikzpicture}
%   }
% \end{frame}
\begin{frame}{Problem}

  \def\arrow{
    (0,0) -- ++(1,0)
    -- ++(0,1) -- ++(0.3, 0)
    -- ++(-0.8, 1) -- ++(-0.8, -1) -- ++(0.3, 0) -- ++(0, -1) }

  \begin{tikzpicture}
    \node[anchor=south west,inner sep=0] (image) at (0,0)
    {\includegraphics[width=\textwidth]{graphics/0000000061_bevdown.png}};
    \begin{scope}[x={(image.south east)},y={(image.north west)}]
      \fill[color=red,scale=0.05,shift={(10,11)},rotate=180,]\arrow;
    \end{scope}
  \end{tikzpicture}
\end{frame}