### 2015-05-03 presentation/pointTrackEnergyWithOccFigure.tex

\begin{tikzpicture}[x=0.8cm,y=0.8cm,3d/perspective eye={4,4,-10},ptstyle/.style={fill=#1,draw,circle,inner sep=1.5pt}]
    \input{Source/twocuboids.tex}
    \draw [->, very thick, draw=blue] (pta4) --  +(90:1) node [anchor=south,align=center] {$a^1(\lambda) = 0$\\$a^2(\lambda) = 1$};
    \draw [->, very thick, draw=red!50!blue] (pta-8) --  +(170:1.4) node [anchor=south,align=center] {$a^1(\lambda) = 0.5$\\$a^2(\lambda) = 0.5$};
    \draw [->, very thick, draw=red] ($(ptb-8)+(.5,-.8)$) --  +(-60:.8) node [anchor=north,align=center] {$a^1(\lambda) = 1$\\$a^2(\lambda) = 0$};
   \begin{scope}[shift={(4,-2)}]
     \path [fill=blue!20,draw] (0,0) ellipse (1.2 and 0.7);
     \path [fill=green!20,draw] (3,0) ellipse (1.2 and 0.7);
     \coordinate (o) at (-2,-1);
     \draw(o) edge [->,very thick] node (xa) {} +(6, 0);
     \node at ($(xa) + (0, -0.2)$) {Depth from camera ($\lambda$)};
     \draw(o) edge [->,very thick] node (ya) {} +(0, 2);
     \node [rotate=90]at ($(ya) + (-0.2, 0)$) {$a^{(1j)}(\lambda)$};
\draw [thick,red] plot [smooth] coordinates {($(o)+(0,1.8)$) ($(o)+(0.2,1.8)$) ($(o) + (0.8, 1.7)$) ($(o)+(1.4,0.3)$) ($(o)+(3.8,0.2)$)($(o)+(4.4,0.1)$) ($(o)+(6.1,0.1)$)};
   \end{scope}
\end{tikzpicture}