### 2015-05-04 CVPR/quals2report.tex

\documentclass[10pt,letterpaper]{report}

\usepackage[margin=1in]{geometry}
%\usepackage{iccv}
\usepackage{booktabs}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{subcaption}

\usepackage{pbox}
\usepackage{verbatim}
\usepackage{bm}
\usepackage{microtype}
\usepackage{url}
\usepackage{multirow}
\usepackage{tikz}
\usetikzlibrary{shadows,trees}

\newlength{\eqs}
\setlength{\eqs}{-0.04in}

\usepackage{enumitem}
\setitemize[0]{leftmargin=15pt}

\newenvironment{tight_itemize}{
\begin{itemize}[leftmargin=10pt]
  \setlength{\topsep}{0pt}
  \setlength{\itemsep}{0pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{itemize}}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

%\newcommand{\bm}{\mathbf{}}
\input{preamble}
\DeclareMathOperator*{\argmin}{\arg\min}

\begin{document}

%%%%%%%%% TITLE
\title{Continous occlusion models and traffic scene understanding}
\author{Vikas Dhiman\\
  {\tt\small dhiman@umich.edu}
  \and
  Mentor: Manmohan Chandrakar\\
  {\tt\small manu@nec-labs.com}
  \and
  Advisor: Jason J Corso\\
  {\tt\small jjcorso@umich.edu}
}
\date{\parbox{\linewidth}{\centering%
  \today\endgraf\bigskip
  Quals 2 report}}
\maketitle
%%%%%%%%% ABSTRACT
\input{Source/abstract.tex}

\tableofcontents

\chapter{Introduction}
\label{sec:intro}
\input{Source/intro}

% % This work is a triple dip: 
% (1) PGM course (2) Qualifier project (3) ICCV submiission

% 
% % Message to be delivered 
% 1. To show the work done
% 2. Exemplify technical powress 
% 3. understanding of graphical models
% 4. Lessons learned: 
%   4.a) occlusion modeling doesn't help much
%   4.b) GPS/Map/Lane are useful
%   4.c) Particle Belief propagation has tough competition from fminunc

% % Course point of view (PGM)
% What is the graphical model that we used for the problem
% What inference method we used for the problem.
% How did the results match to naive realization

% What I did
% GPS/Map/lane orientatio
% Occlusion modeling
% Particle Belief propagation

% What did I learn
% GPS is much more accurate than we would like to think
% Occlusion modeling doesn't help much as we think
% PBP has a surprising competitor

% Outline
% Abstract
% 1 Introduction
% Figure explaining the pipeline?

% 1.5 Related work
\chapter{Related Work}
\label{sec:related}
\input{Source/related}

%
% 2 Modeling
\chapter{Overview of Problem and Our Approach}
\label{sec:setup}
\input{Source/setup}
% Figure explaining the model within a car
% Figure explaining the model within a frame
% Figure explaining the model multiples frames
% 2.1 GPS and Lane information
% 2.2 Occlusion modeling
\chapter{Occlusion-Aware 3D Object Modeling}
\label{sec:TPmodel}
\input{Source/occlusion}
% 2.3 Point tracks
% 2.4 Object detections
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Continuous Energies for Object Localization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{Source/localize}

%
% 3. Inference
% 3.1 fminunc
% 3.2 Particle belief propagation
% 3.3 MCMC
\chapter{Inference}
\input{Source/inference}

%
% 4. Experments and Results
% 4.1 Prove that PGM is useful
% 4.2 Prove that Adding GPS/Lane information is useful
% 4.3 Prove that Adding Point tracks energy is useful
% 4.4 Prove that Adding Occlusion model is useful/not useful
\chapter{Experiments}
\input{Source/experiments}
% 
% 5. Conclusion and Discussions
% 6. Pacify negative results
\chapter{Discussions}
\input{Source/discussions}

\chapter{Appendix}
\label{ref:appendix}
\input{../Paper/suplementarymaterial.tex}

{\small
  \bibliographystyle{ieee}
  \bibliography{continuousocclusion}
}
\end{document}