### 2015-05-11 CVPR/quals2report.tex
Latest version of presentation. Could not add fading because it makes my latex to crash with segmentation fault.

\documentclass[letterpaper]{report}

\usepackage[margin=1in]{geometry}
%\usepackage{iccv}
\usepackage{booktabs}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{subcaption}

\usepackage{pbox}
\usepackage{verbatim}
\usepackage{bm}
\usepackage{microtype}
\usepackage{url}
\usepackage{multirow}
\usepackage{tikz}
\usetikzlibrary{shadows,trees}

\newlength{\eqs}
\setlength{\eqs}{-0.04in}

\usepackage{enumitem}
\setitemize[0]{leftmargin=15pt}

\newenvironment{tight_itemize}{
\begin{itemize}[leftmargin=10pt]
  \setlength{\topsep}{0pt}
  \setlength{\itemsep}{0pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{itemize}}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

%\newcommand{\bm}{\mathbf{}}
\input{preamble}
\DeclareMathOperator*{\argmin}{\arg\min}

\newcommand{\EnergyBBox}{{\pEnergy{bbox}}}
\newcommand{\EnergyTrack}{{ \pEnergy{track}}}
\newcommand{\EnergyTrackNoOcc}{{\pEnergy{trackNoOcc}}}
\newcommand{\EnergyLane}{{\Energy{lane}}}
\newcommand{\EnergySize}{{\Energy{size}}}
\newcommand{\EnergyDyn}{{\Energy{dyn}}}
\newcommand{\EnergyDynHol}{{\Energy{dyn-hol}}}
\newcommand{\EnergyDynOri}{{\Energy{dyn-ori}}}
\newcommand{\EnergyDynVel}{{\Energy{dyn-vel}}}

\newcommand{\xmin}{\xymin{x}}
\newcommand{\xmax}{\xymax{x}}
\newcommand{\ymin}{\xymin{y}}
\newcommand{\ymax}{\xymax{y}}

\begin{document}

%%%%%%%%% TITLE
\title{Continous occlusion models and traffic scene understanding}
\author{Vikas Dhiman\\
  {\tt\small dhiman@umich.edu}
  \and
  Mentor: Manmohan Chandrakar\\
  {\tt\small manu@nec-labs.com}
  \and
  Advisor: Jason J Corso\\
  {\tt\small jjcorso@umich.edu}
}
\date{\parbox{\linewidth}{\centering%
  \today\endgraf\bigskip
  Quals 2 report}}
\maketitle
%%%%%%%%% ABSTRACT
\input{Source/abstract.tex}

\tableofcontents

\chapter{Introduction}
\label{sec:intro}
\input{Source/intro}

% % This work is a triple dip: 
% (1) PGM course (2) Qualifier project (3) ICCV submiission

% 
% % Message to be delivered 
% 1. To show the work done
% 2. Exemplify technical powress 
% 3. understanding of graphical models
% 4. Lessons learned: 
%   4.a) occlusion modeling doesn't help much
%   4.b) GPS/Map/Lane are useful
%   4.c) Particle Belief propagation has tough competition from fminunc

% % Course point of view (PGM)
% What is the graphical model that we used for the problem
% What inference method we used for the problem.
% How did the results match to naive realization

% What I did
% GPS/Map/lane orientatio
% Occlusion modeling
% Particle Belief propagation

% What did I learn
% GPS is much more accurate than we would like to think
% Occlusion modeling doesn't help much as we think
% PBP has a surprising competitor

% Outline
% Abstract
% 1 Introduction
% Figure explaining the pipeline?

% 1.5 Related work
\chapter{Related Work}
\label{sec:related}
\input{Source/related}

%
% 2 Modeling
\chapter{Overview of Problem and Our Approach}
\label{sec:setup}
\input{Source/setup}
% Figure explaining the model within a car
% Figure explaining the model within a frame
% Figure explaining the model multiples frames
% 2.1 GPS and Lane information
% 2.2 Occlusion modeling
\chapter{Occlusion-Aware 3D Object Modeling}
\label{sec:TPmodel}
\input{Source/occlusion}
% 2.3 Point tracks
% 2.4 Object detections
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Continuous Energies for Object Localization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{Source/localize}

%
% 3. Inference
% 3.1 fminunc
% 3.2 Particle belief propagation
% 3.3 MCMC
\chapter{Inference}
\input{Source/inference}

%
% 4. Experments and Results
% 4.1 Prove that PGM is useful
% 4.2 Prove that Adding GPS/Lane information is useful
% 4.3 Prove that Adding Point tracks energy is useful
% 4.4 Prove that Adding Occlusion model is useful/not useful
\chapter{Experiments}
\input{Source/experiments}
% 
% 5. Conclusion and Discussions
% 6. Pacify negative results
\chapter{Discussions}
\input{Source/discussions}

\chapter{Future Work}

We have come up with model of Object detection energy with occlusion, and our plan is evaluate it
and integrate it with the model.

\section{Object detection energy with occlusion}

Object detection is usually followed by non-maximal suppression that results in
discarding similar bounding boxes. When we are jointly optimizing detections
with other cues, it is not usually desirable to go with a single bounding box.
Hence, we keep all the bounding box detections by approximating them with
multi-modal sum of Gaussian like logistic functions. We fit the parametric function of the form 
%
\begin{align}
  S(\bb{i}) = \sum_k A_k \exp(-(\bb{i}-\mu^{(d)}_k)^\top \Sigma^{(d)-1}_k
  (\bb{i}-\mu^{(d)}_k))
\end{align}
%
to detection scores, by non-linear error minimization with initialization from
non-maximal suppressed outputs. Here $\mu^{(d)}_j$ is one of the $k$ modes as a
4D vector representing a single bounding box as $[\minx, \miny, \maxx,
\maxy]^\top$. The optimization is constrained with symmetry and positive
definiteness of $\Sigma^{(d)-1}_k$, $\maxx \ge \minx$ and $\maxy \ge \miny$.

\subsection{Detection scores with occlusion reasoning} 
With our model of $\Ptrans$ described in Section \ref{sec:TPmodel}, we can
compute the probability of a point $\mathbf{u}$ on image be occluded assuming
the point is on TP $i$ with mean depth $\mu^{(i)}_d$ as
\begin{align}
  O_{i}(\mathbf{u}, \mu^{(i)}_d) = 1 - \Ptransmud \enspace .
\end{align}

If we know a portion of our proposed detection bounding box is known to be
occluded, then we would like to decrease the confidence in the detection score
about the localization of that end of the object. Assuming that the occlusion
is often on the boundary of detection bounding boxes, we want to decrease our
confidence on the mean detection boundaries around the occluded boundaries.
%One of the simplest ways will be to scale the appropriate diagonal element of
%$\Sigma_j$ by an appropriate scaling factor proportional to occlusion. But this
%does not model explain the affect of occlusion on the non diagonal terms.
%Hence, we choose a covariance addition model where we compute a occlusion
%covariance matrix, that provides a measure of occlusion in each direction.
To re-model our detection scores scaled by continuous occlusion we sample
$O_{i}(\mathbf{u}, \mu^{(i)}_d)$ at the hypothesized detection boundaries from
GMM $S(.)$ and we augment the detection boundary covariance matrix by
$\mathcal{P}_{j} = \rho_{j}\rho_{j}^\top$ where $\rho_{j} = O_{j}(\mathbf{u},
\mu^{(i)}_d)$. The new covariance matrix in detection score is given by 
  $\Sigma'^{(d)}_j = \mathcal{P}_{j} + \Sigma^{(d)}_j$.
%
%\begin{align}
%\end{align}
%
The detection scores GMM with occlusion is given by replacing the covariance
matrix
%
\begin{align}
  S'(\bb{i}) =
  \sum_j A_j \exp(-(\bb{i}-\mu^{(d)}_j)^\top \Sigma'^{(d)-1}_j
  (\bb{i}-\mu^{(d)}_j))
\end{align}

The energy of detection scores is simply take to be the inverse of the detection score.
\begin{align}
  \Energy{detect}(\{ \relp{i}{t} \}_i, \{ \relp{i}{t-1} \}_i, \{\dimsn{i}\}_i ) = \frac{1}{S'(\bb{i})}
\end{align}

\chapter{Appendix}
\label{sec:appendix}
\input{../Paper/suplementarymaterial.tex}

{\small
  \bibliographystyle{ieee}
  \bibliography{continuousocclusion}
}
\end{document}