### 2015-05-11 presentation/inference.tex

\begin{frame}{Inference}
  \begin{itemize}

    \item
      Metropolic Hastings Inference with Gaussian proposal density (MCMC)

    \item
      Block Coordinate Descent with two kinds of blocks
      \begin{itemize}
        \item
          Optimize pose variables only 
        \item
          Optimize dimension variables only 
      \end{itemize}
  \end{itemize}
  \includegraphics[width=0.5\textwidth, trim=4.5cm 7cm 4.5cm 7cm, clip]{graphics/mcmc_vs_coordinate_descent_60.pdf}
  \includegraphics[width=0.5\textwidth, trim=4.5cm 7cm 4.5cm 7cm, clip]{graphics/mcmc_vs_coordinate_descent_192.pdf}
\end{frame}