### 2015-09-18 presentation/bboxoccfigure.tex
Cleaner and working version of repository.

\usetikzlibrary{intersections}
\begin{tikzpicture}[3d/perspective eye={4,4,-10}]
  \begin{scope}[shift={(0.7,0.9)},scale={10/11},facestyle/.style={fill=green!10,draw=green,thick}]
    \acuboid

  \end{scope}

  \draw [draw=green!50!black,very thick] (b) rectangle (h);
  \path  [name path=gbbox] (b) |- (h);
  \draw [name path=cross1,draw=green!50!black,very thick] (b) -- (h) ;
  \draw [name path=cross2,draw=green!50!black,very thick] (d-| b) -- (b -| h) ;
   \coordinate (gh) at (h);
   \coordinate (gb) at (b);
   \coordinate (gdpb) at (d-|b);

  \begin{scope}[opacity=0.6,facestyle/.style={fill=blue!20,draw=blue!20,very thick}]
    \acuboid
  \end{scope}

  \draw [name path=bbbox,draw=blue!50!black,very thick] (b) rectangle (h) ;

   \path [name intersections={of=cross1 and bbbox,by=E}];
   \path [name intersections={of=gbbox and bbbox,by=F}];
   \path [name intersections={of=cross1 and cross2,by=C}];
   \node [anchor=north east] at (gh) {A};
   \node [anchor=north west] at (gdpb) {D};
   \node [anchor=south west] at (gb) {C};
   \node [anchor=south east] at (gb -| gh) {B};
   \node [anchor=south] at (C) {E};
   \fill [red!30] (gh) -- (E) -- (b) -- (F) -- cycle;
   \node [anchor=west] at ($(C) + (1,0)$) {$v_{AB} = \frac{\text{area(\tikz \node [rectangle,inner sep=0,fill=red!30,text=red!30] {0};)}}{\text{area}(\bigtriangleup ABD)}$};
\end{tikzpicture}