### 2015-11-04 CVPR16/Source/setup.tex

\section{Continuous Occlusion Model}
\label{sec:setup}
The problem of segmenting points to objects often comes up in different
computer vision problems especially in multi-body 
reconstruction and localization. Given correspondence points on
consecutive frames, the problem becomes one of finding out which points belong
to which object, following which one can use the consensus of segmented points
to reconstruct or localize the object more accurately. Without additional 
information, this is a very difficult problem as attempted by semantic segmentation 
and motion segmentation methods. However, in the problems like reconstruction and
localization one can use the hypothesis of the algorithm as the feedback. For
example, in this work we use 3D localization hypotheses to fit ellipsoid
models of objects that can aid in better association of points to objects and
further leading to more accurate localization results.