### 2016-01-25 CVPR16/Rebuttal2/Source/setup.tex

\section{Continuous Occlusion Model}
\label{sec:setup}
We want to solve the problem of segmenting point tracks into object so that the
consensus of points can be used draw inferences about the object. In particular
given a set of correspondence points on consecutive frames $(\trackpj{t},
\trackpj{t-1})$, object localizations $\relp{i}{t}$ and object dimensions
$\dimsn{i}$ we want to figure out which points belong to which object. We call
this the association problem. Even if the pose of the object is not immediately
available we propose the use of hypothesized pose of the object as we show in
the localization estimation experiments.

\def\TP{TP}
A common way of modeling objects, especially traffic participants (TP), in road
scene understanding context is to model the TP as cuboids. However modelingTPs
as cuboids fails to model the uncertainty in the pose and dimensions of TP. Moreover, the cuboid model introduces discontinuity in the problem formulation making
optimization difficulty. With this motivation we propose the use of soft
ellipsoids for modeling TPs that leads to continuous occlusion model that 
accounts for uncertainty in TP position and dimensions.