### 2016-06-22 CVPR/Source/figure_introduction.tex
Changes

\begin{center}
  \includegraphics[width=0.8\linewidth]{graphics/figure1.png}
\end{center}